Summary of the things that I learned

### 2024-09-26 100-introduction.tex
Clean up symbol names, add a link to the UniMath repository

>>>>

Note that throughout this document, links are included to \href{https://arnoudvanderleer.github.io/cs-masters-thesis/toc.html}{documentation} of the corresponding formalized material, formalized either by me or by someone else. It refers to the state of \href{https://github.com/UniMath/UniMath/tree/5eb5c8958c4dddd4219f895bf7bc51547395522d}{the UniMath repository} at commit number \texttt{5eb5c8958c4dddd4219f895bf7bc51547395522d}.

### 2024-09-26 200-category-theory.tex
Clean up symbol names, add a link to the UniMath repository

\section{Notation}
For an object $ c $ in a category $ C $, I will write $ c: C $.

For a morphism $ f $ between objects $ c $ and $ c^\prime $ in a category $ C $, I will write $ f: C(c, c^\prime) $ or $ f: c \to c^\prime $.

For composition of morphisms $ f: C(c, d) $ and $ g: C(d, e) $, I will write $ f \cdot g $.

For composition of functors $ F: A \to B $ and $ G: B \to C $, I will write $ F \bullet G $.

@@

One concept in category theory that can be used to describe a lot of limits and adjunctions is that of a universal arrow (see for example \autocite{MacLane}, Part III)
\begin{definition}
  A \iindex{universal arrow} from an object $ c: C $ to a functor $ F: D \to C $ consists of an object $ d: D $ and a morphism $ f: D(c, F(d)) $ such that for every similar pair $ (d^\prime, f^\prime) $, $ f^\prime $ factors uniquely as $ f \cdot F(g) $ for some $ g: C(d, d^\prime) $:
  \begin{center}
    \begin{tikzcd}
      c \arrow[d, "f"'] \arrow[rd, "f^\prime"] &\\
      F(d) \arrow[r, "F(g)"', dashed] & F(d^\prime)
    \end{tikzcd}
  \end{center}
\end{definition}

Alternatively, we can characterize universal arrows by their action on hom-sets:
\begin{lemma}
  Let $ F: C \to D $ be a functor and $ d: D $ an object. An object $ c: C $ and an arrow $ f: D(d, F(c)) $ form a universal arrow from $ d $ to $ F $ if and only if the function
  \[ (g \mapsto f \cdot F(g)) : C(c, x) \cong D(d, F(x)) \]
  is a bijection.

  Conversely, for all $ c: C $ and $ d: D $, every bijection
  \[ C(c, x) \cong D(d, F(x)) \]
  that is natural in $ x $ arises in this way from some universal arrow $ f: D(d, F(c)) $.
\end{lemma}
\begin{proof}
  See \autocite{MacLane}, Chapter III.2, Proposition 1.
\end{proof}

There is also the dual concept of a universal arrow $ (d, f) $ from a functor $ F $ to an object $ c: C $. Its universal property can be summarized in the following diagram:
\begin{center}
  \begin{tikzcd}
    F(d^\prime)\arrow[rd, "f^\prime"']  \arrow[r, "F(g)", dashed] & F(d) \arrow[d, "f"]\\
    & c
  \end{tikzcd}
\end{center}

@@

An alternative characterization (see \autocite{MacLane}, chapter IV.1, Theorem 2) of an adjunction $ L \dashv R $ is as a natural bijection
\[ \varphi: D(L(c), d) \xrightarrow{\sim} C(c, R(d)). \]
Naturality means that for all $ f: C(c^\prime, c) $, $ g: D(d, d^\prime) $ and $ h: D(L(c), d) $,
\[ \varphi(L(f) \cdot h \cdot g) = f \cdot \varphi(h) \cdot R(g). \]

Lastly, one can construct an adjunction using universal arrows. This lends itself particularly well for a formalization, where it is often preferable to have as little `demonstranda' as possible:
\begin{lemma}
  One can construct an adjunction $ (L, R, \eta, \epsilon) $ as above from only the functor $ L: C \to D $ and, for each $ c: C $, a universal arrow $ (R(c), \epsilon_c) $ from $ L $ to $ c $.
\end{lemma}
\begin{proof}
  See \autocite{MacLane}, Chapter IV.1, Theorem 2 (iv).
\end{proof}

@@

\subsection{Exponential objects}
Note that in the category of sets, for all $ X, Y: \SET $, we have a set $ (X \to Y) $. Also, for all $ X, Y, Z $, there is a (natural) bijection
\[ (X \times Y \to Z) \cong (X \to (Y \to Z)) \]
which we can also write as
\[ \SET(X \times Y, Z) \cong \SET(X, (Y \to Z)). \]
In other words, we have functors $ X \mapsto X \times Y $ and $ Z \mapsto (Y \to Z) $, and these two form an adjunction. The following generalizes this
\begin{definition}
  A category $ C $ has \textit{exponential objects}\index{exponential objects} (or \textit{exponentials}) if for all $ c: C $, the functor $ c^\prime \mapsto c^\prime \times c $ has a right adjoint, which we denote $ d \mapsto d^c $.
\end{definition}

\begin{remark}
  It is actually very well possible that a category does not have all exponentials, but it has some objects $ c, d, d^c: C $ with a natural bijection
  \[ C(d^\prime \times c, d) \cong C(d^\prime, d^c). \]
  Then $ d^c $ is still called an exponential object.
\end{remark}

@@

\begin{example}
  Consider the forgetful functor from the category of commutative rings to the category of sets, sending a ring to its underlying set. This has a left adjoint, sending the set $ \{ 1, 2, \dots, n \} $ to the polynomial ring $ \mathbb Z[X_1, \dots, X_n] $, and more generally, sending $ S $ to the polynomial ring $ \mathbb Z[X_s]_{s : S} $. This ring is then called `the free commutative ring on $ S $'. If $ S $ is finite of size $ n $, the ring is also called `the free commutative ring on $ n $ generators'.

@@

However, sometimes, we have a forgetful functor $ F: C \to D $, but we cannot give a free functor on the entire category $ D $. In such a case, we might still talk about free `objects':
\begin{definition}
  Let $ F: C \to D $ be a forgetful functor. Given $ d: D $, the \textit{free object}\index{free!object} on $ d $ is a universal arrow $ (c, f) $ from $ d $ to $ F $.
\end{definition}

\begin{remark}
  By \autocite{MacLane}, Chapter IV.1, Theorem 2 (ii), if we have a free object on every $ d: D $, we can piece these together to get a free functor associated to $ F $.
\end{remark}

\section{Yoneda}
We can embed a category $ C $ fully faithfully into the functor category $ P C = [\op C, \SET] $ as follows (see \autocite{Kashiwara}, Section 1.4):
\begin{definition}\label{def:Yoneda-embedding}
  The \iindex{Yoneda embedding} $ \yo : C \hookrightarrow P C $ is given on objects by $ \yo(c) = C(-, c) $:
  \[ \yo(c)(d) = C(d, c) \quad \text{and} \quad \yo(c)(f)(g) = f \cdot g \]
  for $ d: C $, $ f: C(d, d^\prime) $ and $ g: C(d^\prime, c) $. It sends a morphism $ f: C(c, c^\prime) $ to the natural transformation $ \yo(f): C(-, c) \Rightarrow C(-, c^\prime) $ given by
  \[ \yo(f)(d)(g) = g \cdot f \]
  for $ d: C $ and $ g: C(d, c) $.
\end{definition}

Now, this embedding has a couple of properties:
\begin{lemma}
  For any $ A: C $ and $ F : P C $, we have an equivalence $ P C(\yo(A), F) \simeq F(A) $, and this equivalence is natural in $ c $ and $ F $.
\end{lemma}
\begin{proof}
  It sends a natural transformation $ \alpha: P C(\yo(A), F) $ to the element $ \alpha_A(\id A) $. Conversely, it sends an element $ x : P(A) $ to the natural transformation $ \alpha $ given by $ \alpha_B(f) = P(f)(x) $ for all $ f: \yo(A)(B) = C(B, A) $. For more details, see \autocite{Kashiwara}, Proposition 1.4.3.
\end{proof}

@@

\begin{remark}
  Also, for any small category $ C $ (small means that the collection of objects is a type in our type universe. For example: for a fixed universe of types, $ \TYPE $ or $ \SET $ are not small categories, the universe of types is not contained in itself), the presheaf category $ P C $ has exponential objects, given by
  \[ (F^G)(X) = \hom{PC}(\yo(X) \times G, F) \]
  the natural transformations from the product functor of the Yoneda embedding of $ X $ and $ G $, to $ F $ (see \autocite{MacLane-Moerdijk}, Section I.6, Proposition 1).
\end{remark}

\begin{definition}
  Suppose that we have a functor $ F: C \to D $, and $ C $ and $ D $ both have binary products. Then for $ A, B : C $, we have morphisms $ \pi_1: A \times B \to A $ and $ \pi_2 : A \times B \to B $ and applying $ F $ to these yields morphisms from $ F(A \times B) $ to $ F(A) $ and $ F(B) $. We then have a product morphism
  \[ \langle F(\pi_1), F(\pi_2) \rangle: F(A \times B) \to F(A) \times F(B). \]
  Now, if this morphism is an isomorphism for all $ A, B : C $, we say that $ F $ \index{preservation of binary products}\textit{preserves binary products}.

  One can imagine that there exists a similar definition of preservation of limits in general, in which the limit morphism of $ F $ applied to the projections from the limit is an isomorphism for any limit in $ C $.
\end{definition}

\begin{definition}\label{def:exponentials-preservation}
  Suppose that we have a functor $ F: C \to D $ that preserves binary products, and $ C $ and $ D $ have exponential objects. Then for $ A, B : C $, we have natural bijections
  \[ \varphi: C(A^B \times B, A) \xrightarrow \sim C(A^B, A^B) \quad \text{and} \quad \psi: D(F(A^B) \times F(B), F(A)) \xrightarrow \sim D(F(A^B), F(A)^{F(B)}) \]
  This gives a morphism $ F(\varphi^{-1}(\id{A^B})) : D(F(A^B \times B), F(A)) $. Precomposing with the inverse of the isomorphism $ f: F(A^B \times B) \xrightarrow \sim F(A^B) \times F(B) $ and applying $ \psi $ gives
  \[ \psi(f^{-1} \cdot \varphi^{-1}(\id{A^B})) : D(F(A^B), F(A)^{F(B)}). \]
  We say that $ F $ \index{preservation of exponentials}\textit{preserves exponentials} if this morphism is an isomorphism for all $ A, B : C $. A name for a functor that preserves exponentials is a\iindex{cartesian functor}.
\end{definition}

\begin{lemma}
  The Yoneda embedding preserves limits.
\end{lemma}
\begin{proof}
  See \autocite{borceux}, Volume 1, Proposition 2.15.5 for the full proof, but note that for binary products, we have bijections
  \[ \yo(A \times B)(X) \xrightarrow \sim \yo(A)(X) \times \yo(B)(X) \]
  sending $ f : C(X, A \times B) $ to the pair $ (f \cdot \pi_1, f \cdot \pi_2) $ and conversely, sending a pair $ (g_1, g_2) : C(X, A) \times C(X, B) $ to the product morphism $ \langle g_1, g_2 \rangle : \yo(X, A \times B) $. This idea is also the core of the proof about general limits.
\end{proof}

\begin{lemma}\label{lem:Yoneda-preserves-exponentials}
  The Yoneda embedding preserves exponentials.
\end{lemma}
\begin{proof}
  First of all, note that for $ A, B, X : C $, we have a sequence of isomorphisms  (\autocite{stackexchange:yoneda-exponentials})
  \begin{align*}
    \yo(A^B)(X) &\cong P C(\yo(X), \yo(A^B))\\
    &\cong C(X, A^B)\\
    &\cong C(X \times B, A)\\
    &\cong P C(\yo(X \times B), \yo(A))\\
    &\cong P C(\yo(X) \times \yo(B), \yo(A))\\
    &\cong P C(\yo(X), \yo(A)^\yo(B))\\
    &\cong (\yo(A)^\yo(B))(X)
  \end{align*}
  using (once or twice) the Yoneda lemma, fully faithfulness of the Yoneda embedding, the property of the exponential object, and the fact that the Yoneda embedding preserves binary products.

  Some calculating shows that when applying this isomorphism to some $ f: \yo(A^B)(X) $, we get the natural transformation $ h: P C(\yo(X) \times \yo(B), \yo(A)) $ given by
  \[ h_Z(g_1, g_2) = \langle g_1, g_2 \rangle \cdot \varphi^{-1}(f) \]
  for all $ Z: C $ and $ (g_1, g_2) : \yo(X)(Z) \times \yo(B)(Z) $ and with the natural bijection
  \[ \varphi: C(X \times B, A) \xrightarrow \sim C(X, A^B). \]

@@

For a functor between categories $ f: C \to D $, given the Yoneda embeddings $ \yo_C : C \to P C $ and $ \yo_D : D \to P D $ (we will often omit the subscript $ C $ and $ D $), we can create a diagram
\begin{center}
  \begin{tikzcd}
    C \arrow[r, "f"] \arrow[d, hookrightarrow, "\yo"] & D \arrow[d, hookrightarrow, "\yo"]\\
    P C & P D \arrow[l, "\op f_*"]
  \end{tikzcd}
\end{center}
Note that the arrows in this diagram are functors, so objects in a category, instead of elements of a set. Therefore, it does not make sense to talk about `equality' of the morphisms along the different paths, but we rather talk about isomorphism in the functor category $ [C, P C] $. If we have such an isomorphism, we say the diagram `2-commutes':
\begin{lemma}\label{lem:Yoneda-restriction-commutes}
  If $ f: C \to D $ is a fully faithful functor, the diagram above 2-commutes.
\end{lemma}
\begin{proof}
  For $ c, d : C $, since $ f $ is fully faithful, we have isomorphisms of sets, given by
  \[ \yo(c)(d) = C(d, c) \xrightarrow[f_{d, c}]{\sim} D(f(d), f(c)) = \op f_*(\yo(f(c)))(d). \]
  Also, for $ g: C(d^\prime, d) $, the following diagram commutes
  \begin{center}
    \begin{tikzcd}
      \yo(c)(d) \arrow[d, "g \cdot -"] \arrow[r, "f_{d, c}"] & \op f_*(\yo(f(c)))(d) \arrow[d, "f_{d^\prime, d}(g) \cdot -"]\\
      \yo(c)(d^\prime) \arrow[r, "f_{d^\prime, c}"] & \op f_*(\yo(f(c)))(d^\prime)
    \end{tikzcd}
  \end{center}
  so the isomorphism is natural in $ d $ and we have $ \yo(c) \cong \op f_*(\yo(f(c))) $ in $ P C $. Lastly, for $ g: C(c, c^\prime) $ and $ d: C $, the following diagram commutes
  \begin{center}
    \begin{tikzcd}
      \yo(c)(d) \arrow[d, "- \cdot g"] \arrow[r, "f_{d, c}"] & \op f_*(\yo(f(c)))(d) \arrow[d, "- \cdot f_{c, c^\prime}(g)"]\\
      \yo(c^\prime)(d) \arrow[r, "f_{d, c^\prime}"] & \op f_*(\yo(f(c^\prime)))(d)
    \end{tikzcd}
  \end{center}
  so the isomorphism is natural in $ c $ and we have $ \yo \cong f \bullet \yo \bullet \op f_* $ in $ [C, P C] $.
\end{proof}

\section{Fibrations}
Let $ P : E \to B $ be a functor. In this case, we will view this as the category $ E $ `lying over' the category $ B $, with for every point $ b: B $, a slice $ E_b = P^{-1}(B) $ lying `above' $ b $.

\begin{definition}
  A morphism $ f: E(y, z) $ is called \iindex{cartesian} if for all $ g: E(x, z) $ and $ h: B(P(x), P(y)) $ with $ h \cdot P(f) = P(g) $, there exists $ \bar h: E(x, y) $ such that $ P(\bar h) = h $ and $ \bar h \cdot f = g $, like illustrated in the following diagram from \autocite{nlab:grothendieck_fibration}
  \begin{center}
    \begin{tikzcd}[sep=large]
      E \arrow[d, "P"] &x \arrow[rr, "\forall g", bend left] \arrow[r, "\exists! \bar h"', dashed] & y \arrow[r, "f"'] & z\\
      B & P(x) \arrow[r, "\forall h"'] \arrow[rr, "P(g)", bend left] & P(y) \arrow[r, "P(f)"'] & P(z)
    \end{tikzcd}
  \end{center}
\end{definition}

\begin{definition}
  $ P $ is a \iindex{fibration} if for all $ y: E $ and morphisms $ f: B(x, P(y)) $, there exist an object $ \bar x: E $ and a cartesian morphism $ \bar f: E(\bar x, y) $ such that $ P(\bar x) = x $ and $ P(\bar f) = f $:
  \begin{center}
    \begin{tikzcd}
      E \arrow[d, "P"] & \bar x \arrow[r, "\exists \bar f", dashed] & y\\
      B & x \arrow[r, "\forall f"] & P(y)
    \end{tikzcd}
  \end{center}
\end{definition}

\section{(Co)slice categories}
Given an object in a category $ c: C $, the morphisms to and from $ c $ constitute the slice and coslice categories
\begin{definition}
  The \iindex{slice category} $ C \downarrow c $ is the category with as objects the morphisms to $ c $:
  \[ (C \downarrow c)_0 = \sum_{c^\prime: C} C(c^\prime, c). \]
  The morphisms from $ (c^\prime, f) $ to $ (c^{\prime\prime}, f^\prime) $ are the morphisms $ g: c^\prime \to c^{\prime\prime} $ making the following diagram commute.
  \begin{center}
    \begin{tikzcd}
      c^\prime \arrow[rr, "g"] \arrow[rd, "f"'] & & c^{\prime\prime} \arrow[ld, "f^\prime"]\\
      & c
    \end{tikzcd}
  \end{center}
\end{definition}
The \textit{coslice category}\index{slice category!co-} $ c \downarrow C $ is similar, but with the morphisms \textit{from} $ c $ instead of \textit{to} $ c $:
\[ (c \downarrow C)_0 = \sum_{c^\prime: C} C(c, c^\prime). \]

Now, if we have an object in a slice category, we can again look at the slice of the slice category over that object. However, this gives us nothing new:
\begin{lemma}
  Let $ (d, f) $ be an object in the slice category $ (C \downarrow c) $. The slice category $ ((C \downarrow c) \downarrow (d, f)) $ is equivalent to $ (C \downarrow d) $.
\end{lemma}
\begin{proof}
  An object of $ ((e, g), \alpha) : ((C \downarrow c) \downarrow (d, f)) $ is an object $ (e, g) : (C \downarrow c) $, together with a morphism $ \alpha: (C \downarrow c)((e, g), (d, f)) $. That is, a morphism $ \alpha: C(e, d) $ such that the obvious triangle commutes (shown in the diagram below on the far left).

>>>>

\section{Notation}
Throughout this thesis, an object $ X $ in a category $ C $ will be denoted by $ X : C $. A morphism $ f $ between objects $ X $ and $ Y $ in a category $ C $ will be denoted by $ f: C(X, Y) $ or sometimes $ f: X \to Y $. Composition of morphisms $ f: C(X, Y) $ and $ g: C(Y, Z) $ will be denoted by $ f \cdot g $. And composition of functors $ F: A \to B $ and $ G: B \to C $ will be denoted by $ F \bullet G $.

@@

One concept in category theory that can be used to describe a lot of limits and adjunctions is that of a universal arrow (see for example \autocite{MacLane}, Part III)
\begin{definition}
  A \iindex{universal arrow} from an object $ X: D $ to a functor $ F: C \to D $ consists of an object $ Y: C $ and a morphism $ f: D(X, F(Y)) $ such that for every similar pair $ (Y^\prime, f^\prime) $, $ f^\prime $ factors uniquely as $ f \cdot F(g) $ for some $ g: C(Y, Y^\prime) $:
  \begin{center}
    \begin{tikzcd}
      X \arrow[d, "f"'] \arrow[rd, "f^\prime"] &\\
      F(Y) \arrow[r, "F(g)"', dashed] & F(Y^\prime)
    \end{tikzcd}
  \end{center}
\end{definition}

Alternatively, we can characterize universal arrows by their action on hom-sets:
\begin{lemma}
  Let $ F: C \to D $ be a functor and $ X: D $ an object. An object $ Y: C $ and an arrow $ f: D(X, F(Y)) $ form a universal arrow from $ X $ to $ F $ if and only if the function
  \[ (g \mapsto f \cdot F(g)) : C(Y, x) \cong D(X, F(x)) \]
  is a bijection.

  Conversely, for all $ Y: C $ and $ X: D $, every bijection
  \[ C(Y, Z) \cong D(X, F(Z)) \]
  that is natural in $ Z $ arises in this way from some universal arrow $ f: D(X, F(Y)) $.
\end{lemma}
\begin{proof}
  See \autocite{MacLane}, Chapter III.2, Proposition 1.
\end{proof}

There is also the dual concept of a universal arrow $ (X, f) $ from a functor $ F $ to an object $ Y: C $. Its universal property can be summarized in the following diagram:
\begin{center}
  \begin{tikzcd}
    F(X^\prime)\arrow[rd, "f^\prime"']  \arrow[r, "F(g)", dashed] & F(X) \arrow[d, "f"]\\
    & Y
  \end{tikzcd}
\end{center}

@@

An alternative characterization (see \autocite{MacLane}, chapter IV.1, Theorem 2) of an adjunction $ L \dashv R $ is as a natural bijection
\[ \varphi: D(L(X), Y) \xrightarrow{\sim} C(X, R(Y)). \]
Naturality means that for all $ f: C(X^\prime, X) $, $ g: D(Y, Y^\prime) $ and $ h: D(L(X), Y) $,
\[ \varphi(L(f) \cdot h \cdot g) = f \cdot \varphi(h) \cdot R(g). \]

Lastly, one can construct an adjunction using universal arrows. This lends itself particularly well for a formalization, where it is often preferable to have as few `demonstranda' as possible:
\begin{lemma}
  One can construct an adjunction $ (L, R, \eta, \epsilon) $ as above from only the functor $ L: C \to D $ and, for each $ X: C $, a universal arrow $ (R(X), \epsilon_X) $ from $ L $ to $ X $.
\end{lemma}
\begin{proof}
  See \autocite{MacLane}, Chapter IV.1, Theorem 2 (iv).
\end{proof}

@@

\subsection{Exponential objects}
Note that in the category of sets, for all $ X, Y: \SET $, we have a set of functions $ (X \to Y) $. Also, for all $ X, Y, Z $, there is a (natural) bijection
\[ (X \times Y \to Z) \cong (X \to (Y \to Z)) \]
which we can also write as
\[ \SET(X \times Y, Z) \cong \SET(X, (Y \to Z)). \]
In other words, we have functors $ X \mapsto X \times Y $ and $ Z \mapsto (Y \to Z) $, and these two form an adjunction. The following generalizes this
\begin{definition}
  A category $ C $ has \textit{exponential objects}\index{exponential objects} (or \textit{exponentials}) if for all $ X: C $, the functor $ X^\prime \mapsto X^\prime \times X $ has a right adjoint, which we denote $ Y \mapsto Y^X $.
\end{definition}

\begin{remark}
  It is actually very well possible that a category does not have all exponentials, but it has some objects $ X, Y, Y^X: C $ with a bijection
  \[ C(X^\prime \times X, Y) \cong C(X^\prime, Y^X) \]
  that is natural in $ X^\prime $. Then $ Y^X $ is still called an exponential object.
\end{remark}

@@

\begin{example}
  Consider the forgetful functor from the category of commutative rings to the category of sets, sending a ring to its underlying set. This has a left adjoint, sending the set $ \{ 1, 2, \dots, n \} $ to the polynomial ring $ \mathbb Z[X_1, \dots, X_n] $, and more generally, sending $ S $ to the polynomial ring $ \mathbb Z[X_s]_{s : S} $. This ring is then called `the free commutative ring on $ S $'. If $ S $ has $ n $ elements, the ring is also called `the free commutative ring on $ n $ generators'.

@@

However, as with exponential object, sometimes we have a forgetful functor $ F: C \to D $, but we cannot give a free functor on the entire category $ D $. In such a case, we might still talk about free `objects':
\begin{definition}
  Let $ F: C \to D $ be a forgetful functor. Given $ X: D $, the \textit{free object}\index{free!object} on $ X $ is a universal arrow $ (Y, f) $ from $ X $ to $ F $.
\end{definition}

\begin{remark}
  By \autocite{MacLane}, Chapter IV.1, Theorem 2 (ii), if we have a free object on every $ X: D $, we can piece these together to get a free functor associated to $ F $.
\end{remark}

\section{Yoneda}
We can embed a category $ C $ fully faithfully into the functor category $ P C = [\op C, \SET] $ as follows (see \autocite{Kashiwara}, Section 1.4):
\begin{definition}\label{def:Yoneda-embedding}
  The \iindex{Yoneda embedding} $ \yo : C \hookrightarrow P C $ is given on objects by $ \yo(Y) = C(-, Y) $:
  \[ \yo(Y)(X) = C(X, Y) \quad \text{and} \quad \yo(Y)(f)(g) = f \cdot g \]
  for $ X: C $, $ f: C(X, X^\prime) $ and $ g: C(X^\prime, Y) $. It sends a morphism $ f: C(Y, Y^\prime) $ to the natural transformation $ \yo(f): C(-, Y) \Rightarrow C(-, Y^\prime) $ given by
  \[ \yo(f)(X)(g) = g \cdot f \]
  for $ X: C $ and $ g: C(X, Y) $.
\end{definition}

Now, this embedding has a couple of properties:
\begin{lemma}
  For any $ Y: C $ and $ F : P C $, we have an equivalence $ P C(\yo(Y), F) \simeq F(Y) $, and this equivalence is natural in $ Y $ and $ F $.
\end{lemma}
\begin{proof}
  It sends a natural transformation $ \alpha: P C(\yo(Y), F) $ to the element $ \alpha_Y(\id Y) $. Conversely, it sends an element $ X : P(Y) $ to the natural transformation $ \alpha $ given by $ \alpha_X(f) = P(f)(X) $ for all $ f: \yo(Y)(X) = C(X, Y) $. For more details, see \autocite{Kashiwara}, Proposition 1.4.3.
\end{proof}

@@

\begin{remark}
  Also, for any small category $ C $ (small means that the collection of objects is a type in our type universe. For example: for a fixed universe of types, $ \TYPE $ or $ \SET $ are not small categories, the universe of types is not contained in itself), the presheaf category $ P C $ has exponential objects, given by
  \[ (F^G)(X) = PC(\yo(X) \times G, F), \]
  the natural transformations from the product functor of the Yoneda embedding of $ X $ and $ G $, to $ F $ (see \autocite{MacLane-Moerdijk}, Section I.6, Proposition 1).
\end{remark}

\begin{definition}
  Suppose that we have a functor $ F: C \to D $, and $ C $ and $ D $ both have binary products. Then for $ X, Y : C $, we have morphisms $ \pi_1: X \times Y \to X $ and $ \pi_2 : X \times Y \to Y $ and applying $ F $ to these yields morphisms from $ F(X \times Y) $ to $ F(X) $ and $ F(Y) $. We then have a product morphism
  \[ \langle F(\pi_1), F(\pi_2) \rangle: F(X \times Y) \to F(X) \times F(Y). \]
  Now, if this morphism is an isomorphism for all $ X, Y : C $, we say that $ F $ \index{preservation of binary products}\textit{preserves binary products}.

  One can imagine that there exists a similar definition of preservation of limits in general, in which the limit morphism of `$ F $ applied to the projections from the limit' is an isomorphism, for any limit in $ C $.
\end{definition}

\begin{definition}\label{def:exponentials-preservation}
  Suppose that we have a functor $ F: C \to D $ that preserves binary products, and $ C $ and $ D $ have exponential objects. Then for $ X, Y : C $, we have natural bijections
  \[ \varphi: C(X^Y \times Y, X) \xrightarrow \sim C(X^Y, X^Y) \quad \text{and} \quad \psi: D(F(X^Y) \times F(Y), F(X)) \xrightarrow \sim D(F(X^Y), F(X)^{F(Y)}) \]
  This gives a morphism $ F(\varphi^{-1}(\id{X^Y})) : D(F(X^Y \times Y), F(X)) $. Precomposing with the inverse of the isomorphism $ f: F(X^Y \times Y) \xrightarrow \sim F(X^Y) \times F(Y) $ and applying $ \psi $ gives
  \[ \psi(f^{-1} \cdot \varphi^{-1}(\id{X^Y})) : D(F(X^Y), F(X)^{F(Y)}). \]
  We say that $ F $ \index{preservation of exponentials}\textit{preserves exponentials} if this morphism is an isomorphism for all $ X, Y : C $. A name for a functor that preserves exponentials is a \iindex{cartesian functor}.
\end{definition}

\begin{lemma}
  The Yoneda embedding preserves limits.
\end{lemma}
\begin{proof}
  See \autocite{borceux}, Volume 1, Proposition 2.15.5 for the full proof, but note that for binary products, we have bijections
  \[ \yo(X \times Y)(Z) \xrightarrow \sim \yo(X)(Z) \times \yo(Y)(Z) \]
  sending $ f : C(Z, X \times Y) $ to the pair $ (f \cdot \pi_1, f \cdot \pi_2) $ and conversely, sending a pair $ (g_1, g_2) : C(Z, X) \times C(Z, Y) $ to the product morphism $ \langle g_1, g_2 \rangle : \yo(Z, X \times Y) $. This idea is also the core of the proof about general limits.
\end{proof}

\begin{lemma}\label{lem:Yoneda-preserves-exponentials}
  The Yoneda embedding preserves exponentials.
\end{lemma}
\begin{proof}
  First of all, note that for $ X, Y, Z : C $, we have a sequence of isomorphisms  (\autocite{stackexchange:yoneda-exponentials})
  \begin{align*}
    \yo(X^Y)(Z) &\cong P C(\yo(Z), \yo(X^Y))\\
    &\cong C(Z, X^Y)\\
    &\cong C(Z \times Y, X)\\
    &\cong P C(\yo(Z \times Y), \yo(X))\\
    &\cong P C(\yo(Z) \times \yo(Y), \yo(X))\\
    &\cong P C(\yo(Z), \yo(X)^\yo(Y))\\
    &\cong (\yo(X)^\yo(Y))(Z)
  \end{align*}
  using (once or twice) the Yoneda lemma, fully faithfulness of the Yoneda embedding, the property of the exponential object, and the fact that the Yoneda embedding preserves binary products.

  Some calculating shows that when applying this isomorphism to some $ f: \yo(X^Y)(Z) $, we get the natural transformation $ h: P C(\yo(Z) \times \yo(Y), \yo(X)) $ given by
  \[ h_Z(g_1, g_2) = \langle g_1, g_2 \rangle \cdot \varphi^{-1}(f) \]
  for all $ Z: C $ and $ (g_1, g_2) : \yo(Z)(Z) \times \yo(Y)(Z) $ and with the natural bijection
  \[ \varphi: C(Z \times Y, X) \xrightarrow \sim C(Z, X^Y). \]

@@

For a functor between categories $ f: C \to D $, given the Yoneda embeddings $ \yo_C : C \to P C $ and $ \yo_D : D \to P D $ (we will often omit the subscript $ C $ and $ D $), we can create a diagram
\begin{center}
  \begin{tikzcd}
    C \arrow[r, "f"] \arrow[d, hookrightarrow, "\yo"] & D \arrow[d, hookrightarrow, "\yo"]\\
    P C & P D \arrow[l, "\op f_*"]
  \end{tikzcd}
\end{center}
Note that the arrows in this diagram are functors, so objects in a category, instead of elements of a set. Therefore, it often does not make a lot of sense to talk about `equality' of the functors along the different paths, but we rather talk about isomorphism in the functor category $ [C, P C] $. If we have such an isomorphism, we say the diagram `2-commutes':
\begin{lemma}\label{lem:Yoneda-restriction-commutes}
  If $ f: C \to D $ is a fully faithful functor, the diagram above 2-commutes.
\end{lemma}
\begin{proof}
  For $ Y, X : C $, since $ f $ is fully faithful, we have isomorphisms of sets, given by
  \[ \yo(Y)(X) = C(X, Y) \xrightarrow[f_{X, Y}]{\sim} D(f(X), f(Y)) = \op f_*(\yo(f(Y)))(X). \]
  Also, for $ g: C(X^\prime, X) $, the following diagram commutes
  \begin{center}
    \begin{tikzcd}
      \yo(Y)(X) \arrow[d, "g \cdot -"] \arrow[r, "f_{X, Y}"] & \op f_*(\yo(f(Y)))(X) \arrow[d, "f_{X^\prime, X}(g) \cdot -"]\\
      \yo(Y)(X^\prime) \arrow[r, "f_{X^\prime, Y}"] & \op f_*(\yo(f(Y)))(X^\prime)
    \end{tikzcd}
  \end{center}
  so the isomorphism is natural in $ X $ and we have $ \yo(Y) \cong \op f_*(\yo(f(Y))) $ in $ P C $. Lastly, for $ g: C(Y, Y^\prime) $ and $ X: C $, the following diagram commutes
  \begin{center}
    \begin{tikzcd}
      \yo(Y)(X) \arrow[d, "- \cdot g"] \arrow[r, "f_{X, Y}"] & \op f_*(\yo(f(Y)))(X) \arrow[d, "- \cdot f_{Y, Y^\prime}(g)"]\\
      \yo(Y^\prime)(X) \arrow[r, "f_{X, Y^\prime}"] & \op f_*(\yo(f(Y^\prime)))(X)
    \end{tikzcd}
  \end{center}
  so the isomorphism is natural in $ Y $ and we have $ \yo \cong f \bullet \yo \bullet \op f_* $ in $ [C, P C] $.
\end{proof}

\section{Fibrations}
Let $ P : E \to B $ be a functor. In this case, we will view this as the category $ E $ `lying over' the category $ B $, with for every point $ X: B $, a slice $ E_X = P^{-1}(B) $ lying `above' $ X $.

\begin{definition}
  A morphism $ f: E(Y, Z) $ is called \iindex{cartesian} if for all $ g: E(X, Z) $ and $ h: B(P(X), P(Y)) $ with $ h \cdot P(f) = P(g) $, there exists $ \bar h: E(X, Y) $ such that $ P(\bar h) = h $ and $ \bar h \cdot f = g $, like illustrated in the following diagram from \autocite{nlab:grothendieck_fibration}
  \begin{center}
    \begin{tikzcd}[sep=large]
      E \arrow[d, "P"] &X \arrow[rr, "\forall g", bend left] \arrow[r, "\exists! \bar h"', dashed] & Y \arrow[r, "f"'] & Z\\
      B & P(X) \arrow[r, "\forall h"'] \arrow[rr, "P(g)", bend left] & P(Y) \arrow[r, "P(f)"'] & P(Z)
    \end{tikzcd}
  \end{center}
\end{definition}

\begin{definition}
  $ P $ is a \iindex{fibration} if for all $ Y: E $ and morphisms $ f: B(X, P(Y)) $, there exist an object $ \bar X: E $ and a cartesian morphism $ \bar f: E(\bar X, Y) $ such that $ P(\bar X) = X $ and $ P(\bar f) = f $:
  \begin{center}
    \begin{tikzcd}
      E \arrow[d, "P"] & \bar X \arrow[r, "\exists \bar f", dashed] & Y\\
      B & X \arrow[r, "\forall f"] & P(Y)
    \end{tikzcd}
  \end{center}
\end{definition}

\section{(Co)slice categories}
Given an object in a category $ X: C $, the morphisms to and from $ X $ constitute the slice and coslice categories
\begin{definition}
  The \iindex{slice category} $ C \downarrow X $ is the category with as objects the morphisms to $ X $:
  \[ (C \downarrow X)_0 = \sum_{Y: C} C(Y, X). \]
  The morphisms from $ (Y_1, f_1) $ to $ (Y_2, f_2) $ are the morphisms $ g: Y_1 \to Y_2 $ making the following diagram commute.
  \begin{center}
    \begin{tikzcd}
      Y_1 \arrow[rr, "g"] \arrow[rd, "f_1"'] & & Y_2 \arrow[ld, "f_2"]\\
      & X
    \end{tikzcd}
  \end{center}
\end{definition}
The \textit{coslice category}\index{slice category!co-} $ X \downarrow C $ is similar, but with the morphisms \textit{from} $ X $ instead of \textit{to} $ X $:
\[ (X \downarrow C)_0 = \sum_{Y: C} C(X, Y). \]
